---
title: "LOND Summary of Mediators GTExV8"
subtitle: 'Audrey Fu Lab'
author: "Jarred Kvamme"
date: '4/8/2021'
output:
  pdf_document: default
  html_document: default
  word_document: default
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/AL_genetabV2.R")

write.path.lond="C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/Tables LOND/"

loadRData <- function(fileName=NULL){
  #loads an RData file, and returns it
  load(fileName)
  get(ls()[ls() != "fileName"])
}


df=loadRData(fileName="C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/loaded_AL_datatables.Rdata")

inter.med=intersect(df$lm1t1$Cis.Gene.ID, df$lm1t2$Trans.Gene.ID)
trans.med=df$lm1t2[-c(match(inter.med, df$lm1t2$Trans.Gene.ID)), ]$Trans.Gene.ID
cis.med=df$lm1t1[-c(match(inter.med, df$lm1t1$Cis.Gene.ID)), ]$Cis.Gene.ID
total.med=length(inter.med)+length(trans.med)+length(cis.med)

summary1=cbind.data.frame(c(length(cis.med), length(trans.med),length(inter.med)), 
                          c(length(cis.med)/total.med, length(trans.med)/total.med, length(inter.med)/total.med))
colnames(summary1)=c("Total.Num.Genes", "Percent.Of.Total")
row.names(summary1)=c("Cis Only","Trans Only","Both Cis & Trans")

library(knitr)
kable(summary1, caption = "Total number of Cis and Trans genes identified as Mediators under MRPC-LOND")
write.csv(summary1, file = paste0(write.path.lond, "Summary.Mediators.csv"), row.names = FALSE)

unq.trans.med=unique(trans.med)
unq.cis.med=unique(cis.med)
unq.inter.med=unique(inter.med)
unq.total.med=length(unq.inter.med)+length(unq.trans.med)+length(unq.cis.med)

summary2=cbind.data.frame(c(length(unq.cis.med), length(unq.trans.med),length(unq.inter.med)), 
                          c(length(unq.cis.med)/unq.total.med, length(unq.trans.med)/unq.total.med, length(unq.inter.med)/unq.total.med))
colnames(summary2)=c("Total.Unique.Genes", "Percent.Of.Total")
row.names(summary2)=c("Cis Only Unique", "Trans Only Unique","Both Cis & Trans")

library(knitr)
kable(summary2, caption = "Total number of Unqiue Cis and Trans genes identified as Mediators under MRPC-LOND")
write.csv(summary2, file = paste0(write.path.lond, "Summary.Mediators.unique.csv"), row.names = FALSE)

```




```{r, echo=FALSE, message=FALSE, warning=FALSE}

#counts and bins for 4 subsets
LTasM=binit(df$lm1t2, target=4)
LCasM=binit(df$lm1t1, target=3)



#Helper plotting function
plotit=function(data=NULL, name=NULL){

  H1=hist(data$counts[,2], breaks = 15, xlab = "# of Tissues Shared", ylab ="# of Genes",
       main = paste("Frequency of Shared Tissues for",name,sep = " "))
  
}

#plot the histograms for 4 subsets
par(mfrow=c(1,2))
plotit(LTasM, "Trans Mediators")
plotit(LCasM, "Cis Mediators")
par(mfrow=c(1,1))

```




```{r, echo=FALSE, message=FALSE, warning=FALSE}

#===============================================================================
sort.bins=function(bins.table=NULL){
  #tissue.names=read.csv("C:/Users/Bruin/Documents/GitHub/MRPC_support/ADDIS_ReRun/tissuenames.csv")
  G=sort(bins.table[,2], index.return=TRUE)
  bins.table=bins.table[G$ix,]
  
  return(bins.table)
}

#===============================================================================

A=sort.bins(LTasM$bins)
B=sort.bins(LCasM$bins)
kable(A, caption="LOND Trans Mediator Tissue Counts", row.names = FALSE)
kable(B, caption="LOND Cis Mediator Tissue Counts", row.names = FALSE)

```



```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Helper function
match.rowN=function(df=NULL, target.vec=NULL){
  
  idx=rep(0, length(target.vec))
  
  for(i in 1:length(target.vec)){
    
    idx[i]=match(target.vec[i], df)
    
  }
  
  return(idx)
  
}

#percentage of unique cis and trans mediator gene types for ADDIS

IX.cis=match.rowN(df=df$lm1t1$Cis.Gene.ID, target.vec = unq.cis.med)
lm1t1.unique=df$lm1t1[IX.cis,]

IX.trans=match.rowN(df=df$lm1t2$Trans.Gene.ID, target.vec = unq.trans.med)
lm1t2.unique=df$lm1t2[IX.trans,]

IX.inter1=match.rowN(df=df$lm1t2$Trans.Gene.ID, target.vec = inter.med)
IX.inter2=match.rowN(df=df$lm1t1$Cis.Gene.ID, target.vec = inter.med)
inter.unique.trans=df$lm1t2[IX.inter1,]
inter.unique.cis=df$lm1t1[IX.inter2,]

lm1t1.unique$cis.Gene.type=as.factor(lm1t1.unique$cis.Gene.type)
lm1t2.unique$trans.Gene.type=as.factor(lm1t2.unique$trans.Gene.type)
inter.unique.cis$cis.Gene.type=as.factor(inter.unique.cis$cis.Gene.type)
inter.unique.trans$trans.Gene.type=as.factor(inter.unique.trans$trans.Gene.type)

percent.both.types=as.data.frame(matrix(0, nrow=9, ncol = 4))
colnames(percent.both.types)=c("Cis.Gene.Type", "Percent.Cis", "Trans.Gene.Type", "Percent.Trans")

percent.both.types$Cis.Gene.Type=c(names(summary(inter.unique.cis$cis.Gene.type)))
percent.both.types$Trans.Gene.Type=names(summary(inter.unique.trans$trans.Gene.type))
percent.both.types$Percent.Cis=c(round(summary(inter.unique.cis$cis.Gene.type)/sum(summary(inter.unique.cis$cis.Gene.type)), 5))
percent.both.types$Percent.Trans=round(summary(inter.unique.trans$trans.Gene.type)/sum(summary(inter.unique.trans$trans.Gene.type)), 5)

percent.LCasM.types=summary(lm1t1.unique$cis.Gene.type)/sum(summary(lm1t1.unique$cis.Gene.type))
percent.LTasM.types=summary(lm1t2.unique$trans.Gene.type)/sum(summary(lm1t2.unique$trans.Gene.type))

kable(percent.LCasM.types, col.names="Percentage", caption="LOND Unique Cis Mediator Gene Types", digits = 5)
kable(percent.LTasM.types, col.names="Percentage", caption="LOND Unique Trans Mediator Gene Types", digits = 5)
kable(percent.both.types[,c(1,2)], caption="LOND gene types for genes found as both Cis and Trans Mediators",
      col.names=c("Gene.Type", "Percent"),row.names = FALSE)

write.csv(percent.LCasM.types, file = paste0(write.path.lond, "percent.unq.LCasM.types.csv"), row.names = FALSE)
write.csv(percent.LTasM.types, file = paste0(write.path.lond, "percent.unq.LTasM.types.csv"), row.names = FALSE)
write.csv(percent.both.types[,c(1,2)], file = paste0(write.path.lond, "percent.both.types.csv"), row.names = FALSE)



```










```{r, echo=FALSE, message=FALSE, warning=FALSE}

BM=read.table(file="C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/mart_export_merged_lncRNA_fixed.txt", sep="\t", header=T)

types4table=function(input=NULL){
  
  idx=as.data.frame(matrix(0, nrow=3, ncol = length(input)))
  test.strs=c("pseudogene", "protein_coding", "lncRNA")
  
  
    
  for(i in 1:length(test.strs)){
    
    print(i)
  
    for(j in 1:length(input)){
      
      idx[i,j]=ifelse(isTRUE(grepl(test.strs[i], input[j], fixed = TRUE)), j, NA)
      
    }
  }
  
  return(idx)
  
}

types_lm1t1=types4table(lm1t1.unique$cis.Gene.type)
types_lm1t2=types4table(lm1t2.unique$trans.Gene.type)
types_wholegenome=types4table(BM$Gene.type)

process.table=function(idx.table=NULL){
  
  test.strs=c("pseudogene", "protein_coding", "lncRNA")
  N=dim(idx.table)[2]
  
  pseudogenes=N-sum(is.na(idx.table[1,]))
  protein_code=N-sum(is.na(idx.table[2,]))
  lncRNAs=N-sum(is.na(idx.table[3,]))
  others=N-(pseudogenes+protein_code+lncRNAs)
  
  counts=c(pseudogenes, protein_code, lncRNAs, others)
  props=counts/N
  
  table1=cbind.data.frame(c(test.strs,"others"),counts, props)
  colnames(table1)=c("Type","Count", "Proportion")
  
  return(table1)
  
}

table.lm1t1=process.table(types_lm1t1)
table.lm1t2=process.table(types_lm1t2)
table.wholegenome=process.table(types_wholegenome)

kable(table.lm1t1, caption = "LOND Unique cis mediator gene types after grouping all gene types into 4 categories: pseudogene, protein_coding, lncRNA, and Others")

kable(table.lm1t2, caption = "LOND Unique trans mediator gene types after grouping all gene types into 4 categories: pseudogene, protein_coding, lncRNA, and Others")

kable(table.wholegenome, caption = "Gene types in the entire genome after grouping all gene types into 4 categories: pseudogene, protein_coding, lncRNA, and Others")
write.csv(table.lm1t1, file = paste0(write.path.lond, "lm1t1.types.4cats.csv"), row.names = FALSE)
write.csv(table.lm1t2, file = paste0(write.path.lond, "lm1t2.types.4cats.csv"), row.names = FALSE)
write.csv(table.wholegenome, file = paste0(write.path.lond, "whole.genome.types.4cats.csv"), row.names = FALSE)

```









\section*{Analysis of Gene Type}
\subsection*{Detecting Pseudogene Enrichment}

To analyze the mediator genes classified under MRPC-LOND, we partitioned the Cis and Trans mediator genes into subsets by their gene types (2 levels: Pseudogene or Non-pseudogene). We then employed hypothesis testing to determine if there was dependence among a mediator gene's position (2 levels: Cis or Trans) and its gene type. We conducted a Chi-Squared test of Independence with the null and alternative hypotheses:

\[ H_0: \text{Gene type is independent of mediator position}  \]
\[ H_A: \text{Gene type is not independent of mediator position}  \]

The resulting $2\times 2$ contingency table is given in \textbf{table 11 and 12}. The test yielded a $\chi_{(1)}^2 = 50.45, \ p \approx 0$ and we therefore rejected the null hypothesis of no dependence. From \textbf{tables 11 and 12} the resulting rejection is due to an approximately $12\%$ difference in the number of pseudogenes between Cis and Trans gene positions. However, under the current test conditions, we are unable to determine if the trans position is $12\%$ enriched with pseudogenes or is the cis position is $12\%$ depleted. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}

#all.addis.df=rbind.data.frame(am1t1.unique, am1t2.unique)

#cis.types=as.factor(all.addis.df$cis.Gene.type)
#trans.types=as.factor(all.addis.df$trans.Gene.type)

cis.types=as.factor(lm1t1.unique$cis.Gene.type)
trans.types=as.factor(lm1t2.unique$trans.Gene.type)

pseudo=function(input=NULL){
  
  replacement=rep(0, length(input))
  test.str="pseudogene"
  
  for(i in 1:length(input)){
    
    replacement[i]=ifelse(isTRUE(grepl(test.str, input[i], fixed = TRUE)), "Pseudogene", "non-pseudo")
    
  }
  
  return(as.factor(replacement))
  
}

#Chi-square test of independence
cis.types=pseudo(cis.types)
cis.types2=cbind.data.frame(cis.types, rep("cis", length(cis.types)))
colnames(cis.types2)=c("type","med")
cis.types2$med=as.factor(cis.types2$med)
trans.types=pseudo(trans.types)
trans.types2=cbind.data.frame(trans.types, rep("trans", length(trans.types)))
colnames(trans.types2)=c("type","med")
trans.types2$med=as.factor(trans.types2$med)

cont.table=table(rbind.data.frame(cis.types2, trans.types2))

chi=chisq.test(as.matrix(cont.table))

cont.table=cbind.data.frame(cont.table[,1], round(cont.table[,1]/sum(cont.table[,1]),4), cont.table[,2], round(cont.table[,2]/sum(cont.table[,2]),4))

cont.table.disp=rbind(cont.table, colSums(cont.table))
cont.table.disp=cbind(cont.table.disp, rowSums(cont.table.disp[,c(1,3)]))

row.names(cont.table.disp)=c("non-pseudo", "pseudo", "row.total")

kable(cont.table.disp, col.names = c("cis",'%cis',"trans",'%trans',"Col.Total"), caption = "2X2 contingency table comparing the cell counts of mediator position and gene_type:pseudogene")
write.csv(cont.table.disp, file = paste0(write.path.lond, "pseudogene.comp.csv"), row.names = FALSE)

# Py=cont.table.disp[3,1]/cont.table.disp[3,5]
# Px=cont.table.disp[1,5]/cont.table.disp[3,5]
# 
# phi.lb=max( -sqrt((Px*Py)/((1-Px)*(1-Py))), -sqrt(((1-Px)*(1-Py))/(Px*Py)) )
# phi.up=min(sqrt( (Px*(1-Py))/ (Py*(1-Px))), sqrt((Py*(1-Px))/(Px*(1-Py))) )
# 
# phi=c(chi$statistic, chi$p.value, chi$parameter, sqrt(chi$statistic/sum(colSums(cont.table))), phi.lb, phi.up)
# names(phi)=c("Chi-Squared", "P", "df", "Phi-coef", "Phi Lower", "Phi Upper")
# kable(phi, caption = "Chi-Squared Test of Independence Summary: gene.type = pseudogene", col.names = "Value")

chi.t=c(chi$statistic, chi$p.value, chi$parameter)
names(chi.t)=c("Chi-Squared", "P", "df")
kable(chi.t, caption = "Chi-Squared Test of Independence Summary: gene.type = pseudogene", col.names = "Value")


```













To further investigate the possibility of trans mediator pseudogene enrichment, we employed a Chi-Squared Goodness of Fit test to determine if enrichment was similar to the proportion of pseudogenes present in the whole genome. In this case, the vector of probabilities ${\bf p}$ is taken to be the observed proportion of pseudo and non-pseudogene types among trans mediator genes, and ${\bf p_0}$ is the expected proportions given by the proportion of pseudo/non-pseudogene types in the entire genome. This leads to the null and alternative hypotheses:

\[ H_0: {\bf p} = {\bf p_0}  \]
\[ H_A: {\bf p} \neq {\bf p_0} \]

The test resulted in a $\chi_{(1)}^2 = 28.72, \ p = 8.345e^{-08}$ and we therefore rejected the null hypothesis of equality of observed and expected proportions. \textbf{Table 13} indicates that the rejection is due to a roughly $5.4\%$ increase of pseudogene gene type among trans mediators relative to the entire genome. 

It is important to note that, because the test statistic in both tests is asymptotically $\chi^2$, the large sample sizes translate into high power and therefore a small difference is translated into a null hypothesis rejection. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Chi-square goodness of fit:
BM=read.table(file="C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/mart_export_merged_lncRNA_fixed.txt", sep="\t", header=T)

all.types=as.factor(BM$Gene.type)
all.types.adj=pseudo(all.types)

observed.probs=summary(all.types.adj)/sum(summary(all.types.adj))

AchsqGOF=chisq.test(summary(trans.types), p=observed.probs)

print(AchsqGOF)

mat1=rbind.data.frame(summary(trans.types), summary(all.types.adj))
per=rbind(summary(trans.types)/sum(summary(trans.types)), summary(all.types.adj)/sum(summary(all.types.adj)))
mat1=cbind.data.frame(mat1[,1], per[,1] ,mat1[,2], per[,2])
colnames(mat1)=c("non-pseudo", '%non-pseudo',"pseudogene", '%pseudogene')
row.names(mat1)=c("Observed","Genome")

kable(mat1, caption = "Chi-Square GOF observed vs. expected proportions of trans gene types ADDIS: gene.type=Pseudogene", digits=4)
write.csv(mat1, file = paste0(write.path.lond, "pseudogene.comp.wg.csv"), row.names = FALSE)

```





\subsection*{Detecting Protein Coding Gene Enrichment}

Similar to the above tests, we also explored the dependence among gene position and gene type with types consisting of protein coding or non-protein coding. Again, we conducted a Chi-square test of independence (\textbf{tables 14 and 15}). The test yielded a $\chi_{(1)}^2 = 175.83, \ p \approx 0$ and we rejected the null hypothesis of no dependence. The rejection is due to an approximately $1:2$ ratio of protein coding to non-protein coding gene types among the cis mediators and an approximately $1:5$ ratio of protein coding to non-protein coding gene types among the trans mediators. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#analysis of protein coding genes

cis.types=as.factor(lm1t1.unique$cis.Gene.type)
trans.types=as.factor(lm1t2.unique$trans.Gene.type)

prot_code=function(input=NULL){
  
  replacement=rep(0, length(input))
  test.str="protein_coding"
  
  for(i in 1:length(input)){
    
    replacement[i]=ifelse(isTRUE(grepl(test.str, input[i], fixed = TRUE)), "protein_coding", "non-protein_coding")
    
  }
  
  return(as.factor(replacement))
  
}

#Chi-square test of independence
cis.types=prot_code(cis.types)
cis.types2=cbind.data.frame(cis.types, rep("cis", length(cis.types)))
colnames(cis.types2)=c("type","med")
cis.types2$med=as.factor(cis.types2$med)
trans.types=prot_code(trans.types)
trans.types2=cbind.data.frame(trans.types, rep("trans", length(trans.types)))
colnames(trans.types2)=c("type","med")
trans.types2$med=as.factor(trans.types2$med)

cont.table2=table(rbind.data.frame(cis.types2, trans.types2))

chi2=chisq.test(as.matrix(cont.table2))

cont.table2=cbind.data.frame(cont.table2[,1], round(cont.table2[,1]/sum(cont.table2[,1]),4), cont.table2[,2], round(cont.table[,2]/sum(cont.table[,2]),4))

cont.table.disp2=rbind(cont.table2, colSums(cont.table2))
cont.table.disp2=cbind(cont.table.disp2, rowSums(cont.table.disp2[,c(1,3)]))

row.names(cont.table.disp2)=c("non-protein_coding", "protein_coding", "row.total")

kable(cont.table.disp2, col.names = c("cis",'%cis',"trans",'%trans',"Col.Total"), caption = "2X2 contingency table comparing the cell counts of mediator position and gene_type:protein coding")
write.csv(cont.table.disp2, file = paste0(write.path.lond, "protein.coding.comp.csv"), row.names = FALSE)
# 
# 
# Py=cont.table.disp2[3,1]/cont.table.disp2[3,5]
# Px=cont.table.disp2[1,5]/cont.table.disp2[3,5]
# 
# phi.lb2=max( -sqrt((Px*Py)/((1-Px)*(1-Py))), -sqrt(((1-Px)*(1-Py))/(Px*Py)) )
# phi.up2=min(sqrt( (Px*(1-Py))/ (Py*(1-Px))), sqrt((Py*(1-Px))/(Px*(1-Py))) )
# 
# phi2=c(chi2$statistic, chi2$p.value, chi2$parameter, sqrt(chi2$statistic/sum(colSums(cont.table2))), phi.lb2, phi.up2)
# names(phi2)=c("Chi-Squared", "P", "df", "Phi-coef", "Phi Lower", "Phi Upper")
# kable(phi2, caption = "Chi-Squared Test of Independence Summary: gene.type=Protein Coding", col.names = "Value")

chi2.t=c(chi2$statistic, chi2$p.value, chi2$parameter)
names(chi2.t)=c("Chi-Squared", "P", "df")
kable(chi2.t, caption = "Chi-Squared Test of Independence Summary: gene.type = protein coding", col.names = "Value")


```




We further explored protein coding gene type enrichment among cis mediators by conducting a Chi-Square GOF test to compare the relative proportion of protein-coding genes among cis mediators to the observed proportion of protein coding genes in the entire genome. The test resulted in a $\chi^2_{(1)} = 39.95, \ p = 2.596e^{-10}$. we therefore rejected the null hypothesis of equality of observed and expected proportions. \textbf{Table 16} indicates that the rejection is due to an approximately $18%$ difference in the proportion of protein coding gene type among cis mediators relative to the entire genome.





```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Chi-square goodness of fit:
BM=read.table(file="C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/mart_export_merged_lncRNA_fixed.txt", sep="\t", header=T)

all.types=as.factor(BM$Gene.type)
all.types.adj=prot_code(all.types)

observed.probs=summary(all.types.adj)/sum(summary(all.types.adj))

AchsqGOF=chisq.test(summary(trans.types), p=observed.probs)

print(AchsqGOF)

mat1=rbind.data.frame(summary(cis.types), summary(all.types.adj))
per=rbind(summary(cis.types)/sum(summary(cis.types)), summary(all.types.adj)/sum(summary(all.types.adj)))
mat1=cbind.data.frame(mat1[,1], per[,1] ,mat1[,2], per[,2])
colnames(mat1)=c("non-protein_coding", '%non-protein_coding',"protein_coding", '%protein_coding')
row.names(mat1)=c("Observed","Genome")

kable(mat1, caption = "Chi-Square GOF observed vs. Expected proportions of cis gene types ADDIS: Type=Protein Coding", digits=4)
write.csv(mat1, file = paste0(write.path.lond, "protein.coding.comp.wg.csv"), row.names = FALSE)

```



\subsection*{Detecting lncRNA Enrichment}

Finally, we explored the dependence among gene position and gene type with type consisting of lncRNA or non-lncRNA using a Chi-square test of independence (\textbf{tables 17 and 18}). The test yielded a $\chi_{(1)}^2 = 15.18, \ p =9.73e^{-05}$, and we rejected the null hypothesis of no dependence. The rejection is due to an approximately $1:4$ ratio of non-lncRNA to lncRNA gene types among the cis mediators and an approximately $1:5$ ratio of non-lncRNA to lncRNA gene types among the trans mediators.   


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#analysis of protein coding genes

cis.types=as.factor(lm1t1.unique$cis.Gene.type)
trans.types=as.factor(lm1t2.unique$trans.Gene.type)

lncrna=function(input=NULL){
  
  replacement=rep(0, length(input))
  test.str="lncRNA"
  
  for(i in 1:length(input)){
    
    replacement[i]=ifelse(isTRUE(grepl(test.str, input[i], fixed = TRUE)), "lncRNA", "non-lncRNA")
    
  }
  
  return(as.factor(replacement))
  
}

#Chi-square test of independence
cis.types=lncrna(cis.types)
cis.types2=cbind.data.frame(cis.types, rep("cis", length(cis.types)))
colnames(cis.types2)=c("type","med")
cis.types2$med=as.factor(cis.types2$med)
trans.types=lncrna(trans.types)
trans.types2=cbind.data.frame(trans.types, rep("trans", length(trans.types)))
colnames(trans.types2)=c("type","med")
trans.types2$med=as.factor(trans.types2$med)

cont.table3=table(rbind.data.frame(cis.types2, trans.types2))

chi3=chisq.test(as.matrix(cont.table3))

cont.table3=cbind.data.frame(cont.table3[,1], round(cont.table3[,1]/sum(cont.table3[,1]),4), cont.table3[,2], round(cont.table3[,2]/sum(cont.table3[,2]),4))

cont.table.disp3=rbind(cont.table3, colSums(cont.table3))
cont.table.disp3=cbind(cont.table.disp3, rowSums(cont.table.disp3[,c(1,3)]))

row.names(cont.table.disp3)=c("non-lncRNA", "lncRNA", "row.total")

kable(cont.table.disp3, col.names = c("cis",'%cis',"trans",'%trans',"Col.Total"), caption = "2X2 contingency table comparing the cell counts of mediator position and gene_type:lncRNA")
write.csv(cont.table.disp3, file = paste0(write.path.lond, "lncRNA.comp.csv"), row.names = FALSE)

# Py=cont.table.disp3[3,1]/cont.table.disp3[3,5]
# Px=cont.table.disp3[1,5]/cont.table.disp3[3,5]
# 
# phi.lb3=max( -sqrt((Px*Py)/((1-Px)*(1-Py))), -sqrt(((1-Px)*(1-Py))/(Px*Py)) )
# phi.up3=min(sqrt( (Px*(1-Py))/ (Py*(1-Px))), sqrt((Py*(1-Px))/(Px*(1-Py))) )
# 
# phi3=c(chi3$statistic, chi3$p.value, chi3$parameter, sqrt(chi3$statistic/sum(colSums(cont.table3))), phi.lb3, phi.up3)
# names(phi3)=c("Chi-Squared", "P", "df", "Phi-coef", "Phi Lower", "Phi Upper")
# kable(phi3, caption = "Chi-Squared Test of Independence Summary: gene.type=lncRNA", col.names = "Value")

chi3.t=c(chi3$statistic, chi3$p.value, chi3$parameter)
names(chi3.t)=c("Chi-Squared", "P", "df")
kable(chi3.t, caption = "Chi-Squared Test of Independence Summary: gene.type = lncRNA", col.names = "Value")

```




To Explore lncRNA gene type enrichment among trans mediators we conducted a Chi-Square GOF test to compare the relative proportion of lncRNA genes among trans mediators to the observed proportion of lncRNA genes in the entire genome. The test resulted in a $\chi^2_{(1)} = 47.96, \ p=4.351e^{-12}$. we therefore rejected the null hypothesis of equality of observed and expected proportions. \textbf{Table 19} indicates that the rejection is due to an approximately $7\%$ difference in the proportion of lncRNA gene type among trans mediators relative to the entire genome.








```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Chi-square goodness of fit:
BM=read.table(file="C:/Users/Bruin/Documents/GitHub/MRPC_support/Reg_Net/mart_export_merged_lncRNA_fixed.txt", sep="\t", header=T)

all.types=as.factor(BM$Gene.type)
all.types.adj=lncrna(all.types)

observed.probs=summary(all.types.adj)/sum(summary(all.types.adj))

AchsqGOF=chisq.test(summary(trans.types), p=observed.probs)

print(AchsqGOF)

mat1=rbind.data.frame(summary(trans.types), summary(all.types.adj))
per=rbind(summary(trans.types)/sum(summary(trans.types)), summary(all.types.adj)/sum(summary(all.types.adj)))
mat1=cbind.data.frame(mat1[,1], per[,1] ,mat1[,2], per[,2])
colnames(mat1)=c("non-lncRNA", '%non-lncRNA',"lncRNA", '%lncRNA')
row.names(mat1)=c("Observed","Genome")

kable(mat1, caption = "Chi-Square GOF observed vs. Expected proportions of gene types ADDIS: Type=lncRNA", digits=4)
write.csv(mat1, file = paste0(write.path.lond, "lncRNA.comp.wg.csv"), row.names = FALSE)


```










































































































































































